---
title: "Credit Card Default"
author: "Dareck Giuliani, Jacob Summerhays, and Sam Wilson"
date: "12/2/2020"
output:
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)

library(tidyverse) #for manipulating and visualizing data
library(MASS) #has function stepAIC(), which allows for easy stepwise regression
library(psych) #dummy code function
library(fastDummies) #dummy code function
library(caret) #for computing cross-validation methods

select <- dplyr::select

options(scipen=999) #effectively eliminates scientific notation

```

# The Data 

This dataset contains information on default payments, demographic factors, credit data, history of payment, and bill statements of credit card clients in Taiwan from April 2005 to September 2005.

## Variables

### Independent Variables

* ID: ID of each client
* LIMIT_BAL: Amount of given credit in NT dollars (includes individual and family/supplementary credit
* SEX: Gender (1=male, 2=female)
* EDUCATION: (1=graduate school, 2=university, 3=high school, 4=others, 5=unknown, 6=unknown)
* MARRIAGE: Marital status (1=married, 2=single, 3=others)
* AGE: Age in years
* PAY_0: Repayment status in September, 2005 (-1=pay duly, 1=payment delay for one month, 2=payment delay for two months, … 8=payment delay for eight months, 9=payment delay for nine months and above)
* PAY_2: Repayment status in August, 2005 (scale same as above)
* PAY_3: Repayment status in July, 2005 (scale same as above)
* PAY_4: Repayment status in June, 2005 (scale same as above)
* PAY_5: Repayment status in May, 2005 (scale same as above)
* PAY_6: Repayment status in April, 2005 (scale same as above)
* BILL_AMT1: Amount of bill statement in September, 2005 (NT dollar)
* BILL_AMT2: Amount of bill statement in August, 2005 (NT dollar)
* BILL_AMT3: Amount of bill statement in July, 2005 (NT dollar)
* BILL_AMT4: Amount of bill statement in June, 2005 (NT dollar)
* BILL_AMT5: Amount of bill statement in May, 2005 (NT dollar)
* BILL_AMT6: Amount of bill statement in April, 2005 (NT dollar)
* PAY_AMT1: Amount of previous payment in September, 2005 (NT dollar)
* PAY_AMT2: Amount of previous payment in August, 2005 (NT dollar)
* PAY_AMT3: Amount of previous payment in July, 2005 (NT dollar)
* PAY_AMT4: Amount of previous payment in June, 2005 (NT dollar)
* PAY_AMT5: Amount of previous payment in May, 2005 (NT dollar)
* PAY_AMT6: Amount of previous payment in April, 2005 (NT dollar)

### Dependent Variables

* default.payment.next.month: Default payment (1=yes, 0=no)

### Exploratory Questions

1. How does the probability of default payment vary by categories of different demographic variables?
2. Which variables are the strongest predictors of default payment?

### Goal:

Create a model that best predicts default rate. 


```{r read in data, echo=TRUE}

#read in data
creditDf <- read.csv('UCI_Credit_Card.csv')
#https://www.kaggle.com/uciml/default-of-credit-card-clients-dataset

#standardize column names
colnames(creditDf) <- tolower(make.names(colnames(creditDf)))

head(creditDf)

```
# Cleaning the Data

```{r understanding the data, echo=TRUE}

#dimensions
dim(creditDf)

#missing values
sum(is.na(creditDf))

#summary
summary(creditDf)
```



```{r correct date types, echo=TRUE}

#make sex a binary 1,0 variables 1=female, 0 = male
creditDf$sex <- if_else(creditDf$sex == 2, true = 1, false = 0)

#make sex a categorical variable
creditDf$sex <- as.factor(creditDf$sex)

#Education contains these three repetitive levels: 4=others, 5=unknown, 6=unknown. This next line is to consolidate the the three levels into 1. 
creditDf$education[creditDf$education == 5] <- 4
creditDf$education[creditDf$education == 6] <- 4

#Some entries have education labeled as 0. 0 is not a known level according to the given key. As such, 0 will be grouped with others and unknown. 
creditDf$education[creditDf$education == 0] <- 4


#Make education a categorical variable
creditDf$education <- as.factor(creditDf$education)

unique(creditDf$education)

#make marriage a categorical variable
creditDf$marriage <- as.factor(creditDf$marriage)

#Some entries of marriage are labeled as 0. However, the key given to us, does not provide meaning for 0. Thus, values for marriage of 0 will be grouped into the other bin. 
creditDf$marriage[creditDf$marriage == 0] <- 3

#make default a categorical variable (1 = yes, 0 = no)
creditDf$default.payment.next.month <- as.factor(creditDf$default.payment.next.month)

#make pay variables into categorical variables
creditDf$pay_0 <- as.factor(creditDf$pay_0)
creditDf$pay_2 <- as.factor(creditDf$pay_2)
creditDf$pay_3 <- as.factor(creditDf$pay_3)
creditDf$pay_4 <- as.factor(creditDf$pay_4)
creditDf$pay_5 <- as.factor(creditDf$pay_5)
creditDf$pay_6 <- as.factor(creditDf$pay_6)

#Create a variable for average bill and for average payment.
creditDf <- creditDf %>%
  mutate(
    bill_avg = round((creditDf$bill_amt1 + creditDf$bill_amt2 + creditDf$bill_amt3 + creditDf$bill_amt4 + creditDf$bill_amt5 + creditDf$bill_amt6)/6, 2),
    pay_avg = round((creditDf$pay_amt1 + creditDf$pay_amt2 +creditDf$pay_amt3 + creditDf$pay_amt4 + creditDf$pay_amt5 + creditDf$pay_amt6 ) /6,2),
    #This creates observations with both + infinity and - infinity. I'm not sure if a regression can handle those values. 
    bill_pay_ratio = round(bill_avg/pay_avg, 2), 
    #Credit scores often judge a person on their credit usage. The closer to the limit a person is, the more of a negative impact on credit score. This metric will divide the average bill amount by the limit balance to get a sense of the average credit usage.
    avg_credit_usage = round(bill_avg/limit_bal, 2)
  )

#When looking through this data, there are a number of customers who have not paid in the last 6 months. This variable is a count of the number of payments in the last 6 months.





#collapse some of the dummy variables

#useful subsets
categorical <- select(creditDf, c(sex, education, marriage, default.payment.next.month, pay_0, pay_2, pay_3, pay_4, pay_5, pay_6))
continuous <- select(creditDf, -c(sex, education, marriage, default.payment.next.month, pay_0, pay_2, pay_3, pay_4, pay_5, pay_6))
dummy <- select(creditDf, c(education, marriage, pay_0, pay_2, pay_3, pay_4, pay_5, pay_6))

#creates dummy variables for all categorical variables with more than two levels
creditDf_Dummy <- dummy_cols(creditDf, select_columns = colnames(dummy))

summary(creditDf)

```

# Statistics

## Descriptive Statistics

## Initial Visualizations

```{r IV vs. DV, echo=TRUE}

#density plots of iV
creditDf %>% 
  select(-c(id)) %>%
  sample_n(50) %>% #samples from the data set to improve speed
  gather(-default.payment.next.month, key = 'iv', value = 'value') %>%
  ggplot(aes(x = value,  fill = default.payment.next.month)) +
  geom_density(position = 'fill', alpha = .5) +
  facet_wrap(~iv, scales = 'free') + 
  ylab('Density') +
  xlab('Independent Variables') +
  theme_minimal() +
  theme(
    #axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = 'bottom',
    legend.title = element_blank(),
    #axis.text.x = element_blank()
  ) +
  scale_color_discrete(labels = c('Default', 'No Default' ))
  

```


```{r IV box plots, echo=TRUE}

creditDf %>% 
  select(-c(id)) %>%
  sample_n(50) %>% #samples from the data set to improve speed
  gather(-default.payment.next.month, key = 'iv', value = 'value') %>%
  ggplot(aes(y = value,  group = default.payment.next.month, x = default.payment.next.month)) +
  geom_boxplot() +
  facet_wrap(~iv, scales = 'free') +
  theme_minimal() +
  theme(
    #axis.text.y = element_blank()
  )


```


```{r variable histograms, echo=TRUE}

#histograms of categorical data
categorical %>%
  sample_n(30) %>%
  gather(key = 'var', value = 'value') %>%
  ggplot(aes(x = value)) +
  geom_histogram(stat = 'count') +
  facet_wrap(~var, scales = 'free') +
  xlab('Variable') +
  ylab('Count') +
  ggtitle('Histograms of All Categorical Variables') +
  theme_minimal() +
  theme(
    #axis.text.x = element_text(angle = 45, hjust = 1),
    #axis.text.x = element_blank(),
    axis.title.x = element_blank()
  )

#histograms of continuous data
continuous %>%
  sample_n(30) %>%
  gather(key = 'var', value = 'value') %>%
  ggplot(aes(x = value)) +
  geom_density() +
  facet_wrap(~var, scales = 'free') +
  ylab('Density') +
  ggtitle('Density Plots of All Continuous Variables') +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    #axis.text.x = element_blank(),
    axis.title.x = element_blank()
  )
  
  

```


```{r correlation with IV, echo=TRUE}

creditDf %>%
  sample_n(30) %>%
  pairs(lower.panel = NULL)
  

```


```{r aScatter Plots, echo=TRUE}

#Average Bill Amount vs. Average Pay Amount
creditDf %>%
  ggplot(aes(x = bill_avg, y = pay_avg, color = default.payment.next.month)) +
  geom_point()

#Age vs. Average Bill Amount
creditDf %>%
  ggplot(aes(x = age, y = pay_avg, color = default.payment.next.month)) +
  geom_point()

creditDf %>%
  ggplot(aes(x = age, y = bill_avg, color = default.payment.next.month)) +
  geom_point()

creditDf %>%
  ggplot(aes(x = avg_credit_usage, fill = default.payment.next.month)) +
  geom_bar()


```


# KNN to segemnt credit card customers



# logistic regression to predict degfault

#Training and Test Data

```{r training test split, echo=TRUE}

#creates a sample of 80% of the data
sample_size <- floor(.8 * nrow(creditDf))

#
set.seed(123)
train_index <- sample(seq_len(nrow(creditDf)), size = sample_size)

#
train <- creditDf_Dummy[train_index, ]
test <- creditDf_Dummy[-train_index, ]

```

# Computing Stepwise Logistic Regression

The stepwise logistic regression can be easily computed using the R function stepAIC() available in the MASS package. It performs model selection by AIC. It has an option called direction, which can have the following values: “both”, “forward”, “backward” (see Chapter @ref(stepwise-regression)).

http://www.sthda.com/english/articles/36-classification-methods-essentials/150-stepwise-logistic-regression-essentials-in-r/

```{r stepwise logistic}

#logit_model1 <- glm(default.payment.next.month ~ . , data = train, family = 'binomial') %>%
#  stepAIC(trace = FALSE)
#After 30 minutes, R continued to process

logit_model1 <- glm(default.payment.next.month ~ limit_bal + sex +  , data = train, family = 'binomial') %>%
#  stepAIC(trace = FALSE)


summary(logit_model1)
```


```{r stepwise logistic continued}



#interaction between gender and marriage. Are make singles more likely to default?

```


#ROC Curve

#Confusion Matrix

# random forrest to predict default 