---
title: "Credit Card Default"
author: "Dareck Giuliani, Jacob Summerhays, and Sam Wilson"
date: "12/2/2020"
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    toc: yes
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)

library(tidyverse)

options(scipen=999)

```

# The Data 

This dataset contains information on default payments, demographic factors, credit data, history of payment, and bill statements of credit card clients in Taiwan from April 2005 to September 2005.

## Variables

### Independent Variables

* ID: ID of each client
* LIMIT_BAL: Amount of given credit in NT dollars (includes individual and family/supplementary credit
* SEX: Gender (1=male, 2=female)
* EDUCATION: (1=graduate school, 2=university, 3=high school, 4=others, 5=unknown, 6=unknown)
* MARRIAGE: Marital status (1=married, 2=single, 3=others)
* AGE: Age in years
* PAY_0: Repayment status in September, 2005 (-1=pay duly, 1=payment delay for one month, 2=payment delay for two months, … 8=payment delay for eight months, 9=payment delay for nine months and above)
* PAY_2: Repayment status in August, 2005 (scale same as above)
* PAY_3: Repayment status in July, 2005 (scale same as above)
* PAY_4: Repayment status in June, 2005 (scale same as above)
* PAY_5: Repayment status in May, 2005 (scale same as above)
* PAY_6: Repayment status in April, 2005 (scale same as above)
* BILL_AMT1: Amount of bill statement in September, 2005 (NT dollar)
* BILL_AMT2: Amount of bill statement in August, 2005 (NT dollar)
* BILL_AMT3: Amount of bill statement in July, 2005 (NT dollar)
* BILL_AMT4: Amount of bill statement in June, 2005 (NT dollar)
* BILL_AMT5: Amount of bill statement in May, 2005 (NT dollar)
* BILL_AMT6: Amount of bill statement in April, 2005 (NT dollar)
* PAY_AMT1: Amount of previous payment in September, 2005 (NT dollar)
* PAY_AMT2: Amount of previous payment in August, 2005 (NT dollar)
* PAY_AMT3: Amount of previous payment in July, 2005 (NT dollar)
* PAY_AMT4: Amount of previous payment in June, 2005 (NT dollar)
* PAY_AMT5: Amount of previous payment in May, 2005 (NT dollar)
* PAY_AMT6: Amount of previous payment in April, 2005 (NT dollar)

### Dependent Variables

* default.payment.next.month: Default payment (1=yes, 0=no)

```{r read in data, echo=TRUE}

#read in data
creditDf <- read.csv('UCI_Credit_Card.csv')

#standardize column names
colnames(creditDf) <- tolower(make.names(colnames(creditDf)))

head(creditDf)

```
# Cleaning the Data

```{r understanding the data, echo=TRUE}

#dimensions
dim(creditDf)

#missing values
sum(is.na(creditDf))

#summary
summary(creditDf)
```



```{r correct date types, echo=TRUE}

#make sex a binary 1,0 variables 1=female, 0 = male
creditDf$sex <- if_else(creditDf$sex == 2, true = 1, false = 0)

#make sex a categorical variable
creditDf$sex <- as.factor(creditDf$sex)


#make education a categorical variable
creditDf$education <- as.factor(creditDf$education)

#make education a dummy variable

#make marriage a categorical variable
creditDf$marriage <- as.factor(creditDf$marriage)

#make default a categorical variable (1 = yes, 0 = no)
creditDf$default.payment.next.month <- as.factor(creditDf$default.payment.next.month)

#make pay variables into categorical variables
creditDf$pay_0 <- as.factor(creditDf$pay_0)
creditDf$pay_2 <- as.factor(creditDf$pay_2)
creditDf$pay_3 <- as.factor(creditDf$pay_3)
creditDf$pay_4 <- as.factor(creditDf$pay_4)
creditDf$pay_5 <- as.factor(creditDf$pay_5)
creditDf$pay_6 <- as.factor(creditDf$pay_6)


#useful subsets
categorical <- select(creditDf, c(sex, education, marriage, default.payment.next.month, pay_0, pay_2, pay_3, pay_4, pay_5, pay_6))
continuous <- select(creditDf, -c(sex, education, marriage, default.payment.next.month, pay_0, pay_2, pay_3, pay_4, pay_5, pay_6))

summary(creditDf)

```

# Statistics

## Descriptive Statistics

## Initial Visualizations

```{r IV vs. DV, echo=TRUE}

#density plots of iV
creditDf %>% 
  select(-c(id)) %>%
  sample_n(50) %>% #samples from the data set to improve speed
  gather(-default.payment.next.month, key = 'iv', value = 'value') %>%
  ggplot(aes(x = value,  fill = default.payment.next.month)) +
  geom_density(position = 'fill', alpha = .5) +
  facet_wrap(~iv, scales = 'free') + 
  ylab('Density') +
  xlab('Independent Variables') +
  theme_minimal() +
  theme(
    #axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = 'bottom',
    legend.title = element_blank(),
    #axis.text.x = element_blank()
  ) +
  scale_color_discrete(labels = c('Default', 'No Default' ))
  

```
```{r IV box plots, echo=TRUE}

creditDf %>% 
  select(-c(id)) %>%
  sample_n(50) %>% #samples from the data set to improve speed
  gather(-default.payment.next.month, key = 'iv', value = 'value') %>%
  ggplot(aes(y = value,  group = default.payment.next.month, x = default.payment.next.month)) +
  geom_boxplot() +
  facet_wrap(~iv, scales = 'free') +
  theme_minimal() +
  theme(
    #axis.text.y = element_blank()
  )


```


```{r variable histograms, echo=TRUE}

#histograms of categorical data
categorical %>%
  sample_n(30) %>%
  gather(key = 'var', value = 'value') %>%
  ggplot(aes(x = value)) +
  geom_histogram(stat = 'count') +
  facet_wrap(~var, scales = 'free') +
  xlab('Variable') +
  ylab('Count') +
  ggtitle('Histograms of All Categorical Variables') +
  theme_minimal() +
  theme(
    #axis.text.x = element_text(angle = 45, hjust = 1),
    #axis.text.x = element_blank(),
    axis.title.x = element_blank()
  )

#histograms of continuous data
continuous %>%
  sample_n(30) %>%
  gather(key = 'var', value = 'value') %>%
  ggplot(aes(x = value)) +
  geom_density() +
  facet_wrap(~var, scales = 'free') +
  ylab('Density') +
  ggtitle('Density Plots of All Continuous Variables') +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    #axis.text.x = element_blank(),
    axis.title.x = element_blank()
  )
  
  

```



```{r correlation with IV, echo=TRUE}

creditDf %>%
  sample_n(30) %>%
  pairs(lower.panel = NULL)
  

```


# KNN to segemnt credit card customers

# logistic regression to predict degfault

# random forrest to predict default 